---
title: "The DART Matrix"
author: "Daniel Spakowicz"
date: "4/24/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(xtable)
library(ggplot2)
```

This is the start of Dan's effort on the final project for CBB750. The intention of this document is to perform some exploratory data analysis and to create summary slides for the presentation.

```{r}
# Read in data 
x <- read.csv("TheDartMatrix-deidentified.csv", as.is = TRUE)

# Read in variable key that defines classes
key <- read.csv("variable_key.csv")
levels(key$class)
# Set classes
factors <- grep("factor", key$class)
char <- grep("character", key$class)
int <- grep("integer", key$class)
num <- grep("numeric", key$class)

x[,factors] <- lapply(x[factors], factor)
x[,char] <- lapply(x[char], as.character)
x[,int] <- lapply(x[int], as.integer)
x[,num] <- lapply(x[num], as.numeric)

```

Let's look at a quick table of when `fits_matrix` is called and when the DART consult is called.

```{r, results="asis", message=FALSE}
table(x$fits_matrix)
print(xtable(table(fits_matrix = x$fits_matrix, consult_dart = x$consult_dart)),
      comment = FALSE)
```

I don't understand why `consult_dart` is 0-3.

Now I'll look at the frequency of the predictor variables.

```{r}
mat <- data.frame(lapply(x[,5:21], as.numeric))
means <- data.frame(lapply(mat, function(x) mean(x-1, na.rm = TRUE)))
tmeans <- data.frame(var = names(means), perc_occurrence = t(means))
ggplot(tmeans, aes(x = reorder(var, perc_occurrence), y = perc_occurrence)) +
  geom_bar(stat = "identity", aes(fill = perc_occurrence)) +
  coord_flip() +
  labs(y = "Percent Occurence",
       x = "Observation") +
  theme_bw() +
  theme(legend.position = "none")+
  ggsave("var_percOccurrence.pdf")
```

